---
title: "Final Project"
author: "Veronica Williamson"
output: html_document
date: "2024-11-07"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#Download packages/libraries
library(mgcv)
library(rstatix)
library(writexl)
library(vcd)
library(ggplot2)
library(dplyr)
library(tidyr)
library(tidyverse)
library(emmeans)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
#Final Project Data. Install and arrange
library(readxl)
og_df <- read_excel("~/Desktop/ARCGIS AND R/final project/Data collection set up.xlsx")
og_df$TreatmentID <- factor(og_df$TreatmentID) 
og_df <- og_df %>%
   mutate(TreatmentID = factor(TreatmentID, 
                              levels = 1:6, 
                              labels = c("Past treatment", 
                                         "Pstrig treatment", 
                                         "Multi-Species Treatment", 
                                         "Single Genotype Treatment", 
                                         "Multi-genet treatment", 
                                         "Single Apal")))
```

# The Effects of Coral Diversity on *Acropora palmata* Growth in a Land Based Nursery

Null Hypothesis: There is no significant difference in area growth among
the 6 treatments over the 12-week period. <br> Alternative Hypothesis:
At least one of the 6 treatments results in a significant difference in
area growth over the 12-week period.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
Avg_df <- og_df %>%
  group_by(Time_period, TreatmentID) %>%
summarize(
    average_area = mean(Area_cmsq, na.rm = TRUE), 
    SEM = sd(Area_cmsq, na.rm = TRUE) / sqrt(n()), 
    .groups = "drop") 
  

ggplot(Avg_df, aes(x = Time_period, y = average_area, group = TreatmentID, color = TreatmentID)) +
   geom_line() +
  labs(title = "Acropora Area (cm^2) Based on Treatment Type Over Three Months", x = "Time Period", y = "Average Area (cm^2)", color = "Treatment Type") +
  theme_minimal() 
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Box Plot to visualize distributions
# **boxplot(og_df**$Area_cmsq ~ og_df$Time_period)
# 
# Calulate the average difference in growth at halfway then end of
# experiment. WHY DID I DO THIS? NOT SURE ITS HELPFUL \*\* halfway \<-
# Avg_df %\>% filter(Time_period %in% c(0, 6)) %\>% select(-SEM) %\>%
# group_by(TreatmentID,Time_period) %\>% pivot_wider(names_from =
# Time_period, values_from = average_area, names_prefix = "Time_period\_")
# %\>% mutate(growth_difference = Time_period_6 - Time_period_0)
# 
# ending \<- Avg_df %\>% filter(Time_period %in% c(0, 12)) %\>%
# select(-SEM) %\>% group_by(TreatmentID,Time_period) %\>%
# pivot_wider(names_from = Time_period, values_from = average_area,
# names_prefix = "Time_period\_") %\>% mutate(growth_difference =
# Time_period_12 - Time_period_0) \*\*

# 
# (If there is an interaction between treatment and time. Null hypothesis
# (H₀) for interaction: There is no significant interaction between
# treatment and time, i.e., the effect of treatment on coral growth is the
# same across all time points. Alternative hypothesis (H₁) for
# interaction: There is a significant interaction between treatment and
# time, i.e., the effect of the treatment on coral growth changes over the
# 12 weeks.)
# 
# Now I want to test whether the means are significantly different from
# each other. For sake of this class I will run an LMM. This wont work for
# some reason


# **LMM_model \<- lmer(Area_cmsq \~ TreatmentID + (1\|Time_period), data =
# og_df)**
# 
# Instead I will run a linear model to get a better idea of what im
# looking for
```

## Analysis

A linear model was run to examine the relationship between coral growth
(Area_cmsq) and two key factors: TreatmentID and Time_period (1-12). The
formula below evaluates the main effects of each factor and the
interaction effect between TreatmentID and Time_period. This allows us
to see whether the effect of treatment on coral growth changes over
time.

```{r}
lm_model <- lm(Area_cmsq ~ TreatmentID * Time_period, data = og_df)
```

## Results

Residual standard error: 2.244 on 456 degrees of freedom <br> Multiple
R-squared: 0.1807 <br> Adjusted R-squared: 0.1609 <br> F-statistic:
9.142 on 11 and 456 DF <br> p-value: 7.562e-15 <br> The results suggest
that treatment effects alone were not significant, but that certain time
periods and specific treatment-time interactions ( Treatment 4: Single
Genotype Treatment at later time periods) had a significant effect on
coral growth.

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
ggplot(og_df, aes(x = Time_period, y = Area_cmsq, color = TreatmentID)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE) +
  labs(title = "Coral Growth Over Time by Treatment",
       x = "Time Period", y = "Growth (cm²)")


```
